# Analysis of Feature sets for Churn reduction
========================================================
Author  : Ravi Sangili
Date    : 01/22/2014

## Data import of the TRaining Features set:

### List of members with memberships renewl due: Nov 2013 - Oct 2014
### Target label : No Churn - Members who renewed
### Target label : Churn    - Members who did not renew within 2 months of expiry


```{r}
if (!require("Hmisc")) {
  install.packages("Hmisc", repos="http://cran.rstudio.com/") 
  library('Hmisc')
}
if (!require("DMwR")) {
  install.packages("DMwR", repos="http://cran.rstudio.com/") 
  library('DMwR')
}

txn <- read.csv("feature for training.csv",header=TRUE, sep=",",stringsAsFactors=FALSE)

```

### Data Cleanup - Identify features with NA
``` {r}
nrow(txn[is.na(txn$CnBio.Age),])

nrow( txn[is.na(txn$Distance),])

nrow( txn[is.na(txn$Visit),])

nrow( txn[is.na(txn$NetAmt),])

nrow( txn[is.na(txn$CnBio.Constit.Code),])

```


#### Remove the record with no zip code or distance

``` {r}


txn <- txn[-is.na(txn$Distance),]

``` 

#### Fill the Attr 5.01 with "no donor" and Attr 4.01 with "no recipient", if blank

``` {r}
mem_donor_blank <- txn[!(grepl("Donor*",txn$CnAttrCat.5.01.Description)),]

mem_donor_blank$CnAttrCat.5.01.Description <- c("no donor")

txn[(!grepl("Donor",txn$CnAttrCat.5.01.Description)), ]  <- mem_donor_blank$CnAttrCat.5.01.Description 

txn[(!grepl("Donor",txn$CnAttrCat.5.01.Description)),c('CnAttrCat.5.01.Description') ]  <- mem_donor_blank$CnAttrCat.5.01.Description 


mem_donor <- txn[grepl("Donor*",txn$CnAttrCat.5.01.Description),]

mem_donor_churn <- mem_donor[mem_donor$CnMem.1.01.Standing == "churn",]

mem_donor_nochurn <- mem_donor[mem_donor$CnMem.1.01.Standing == "no churn",]

donor_recept <- txn[grepl("Recipient*",txn$CnAttrCat.4.01.Description),]

donor_recept_blank <- txn[!grepl("Recipient",txn$CnAttrCat.4.01.Description),]

donor_recept_blank$CnAttrCat.4.01.Description <- c("no recipient")

txn[(!grepl("Recipient",txn$CnAttrCat.4.01.Description)),c('CnAttrCat.4.01.Description') ]  <-  donor_recept_blank$CnAttrCat.4.01.Description 

donor_recept_churn <- donor_recept[donor_recept$CnMem.1.01.Standing == "churn",]

# mem_donor_blank <- txn[!(grepl("Donor*",txn$CnAttrCat.5.01.Description)),]

# mem_donor_blank$CnAttrCat.5.01.Description <- c("no donor")
 

```

### Summary statistics of significant features

``` {r}

summary(txn$Events)
summary(txn$Visits)
summary (txn$NetAmt)
summary (txn$Distance)

```

### Summary statistics for independent variables

``` {r}

boxplot(txn$Events)
hist(txn$CnBio.Age)
hist(txn$Events)
hist(txn$Distance)
```

### Graphs showing relationship between dependent variable and independent variables

``` {r}
histogram(~NetAmt|CnMem.1.01.Standing,txn,na.rm=T)
histogram(~Events|CnMem.1.01.Standing,txn,na.rm=T)
bwplot(CnMem.1.01.Standing~Visits,data=txn)
bwplot(CnMem.1.01.Standing~Visits,data=txn,xlim=1:100)
bwplot(CnMem.1.01.Standing~Events,data=txn)
bwplot(CnMem.1.01.Standing~NetAmt,data=txn,xlim=1:1500)
bwplot(CnMem.1.01.Standing~NetAmt,data=txn,xlim=1:500)
xyplot(Distance ~ Visits |CnMem.1.01.Standing ,txn,ylim=0:100,xlim=0:25)
xyplot(Distance ~ Visits |CnMem.1.01.Standing ,txn)
```


``` {r}
write.csv(feature_df, "C:/Ravi/capstone/clean features.csv")
```
