###Energy Analytics - Assignment 3
#### James Weaver - jw3591 - Feb 2015

*1. Import the data in R and plot the OntarioDemand (column D) and the respective histogram.*

```{r echo = FALSE, message = FALSE, include=FALSE}
library(psych)
library(ggplot2)
library(tseries)
library(stats)

od <- read.csv("Desktop/MSBA/Module 4/Energy Analytics/OntarioDemand.csv")
#od <- read.csv("/Users/jamieweaver/Dropbox//NYU mod4/Energy Analytics/OntarioDemand.csv")
days <- length(unique(od$Date))
od$Date <- as.Date(od$Date, "%d-%b-%y")
```

The time series data consist of `r dim(od)[1]` observarions characterized by `r dim(od)[2]` attributes. Each observation represents 1 of 24 hours in a day and there are `r days` days in the time series range `r min(od$Date)` to `r max(od$Date)`. Each observation includes energy demand information for the total market and for Ontario specifically. Figure 1 displays TIME SERIES PLOT, SDs,

Over this time period, the average hourly energy demand in Ontario was `r format(round(mean(od$Ontario.Demand, 0)), scientific=FALSE)` kWh with standard deviation `r round(sd(od$Ontario.Demand), 0)` and is approximately normally distributed (Figure 2) and confirmed by the Kolmogorov-Smirnov test (p<0.001).

**Figure 1. Average Ontario energy demand per hour, 5/2002-3/2004**
```{r, echo = FALSE, message=FALSE}
od.mean <- aggregate(Ontario.Demand ~ Date, od, mean)
names(od.mean) <- c("Date", "od.mean")
od.sd <- aggregate(Ontario.Demand ~ Date, od, sd)
names(od.sd) <- c("Date", "od.sd")
od.sd <- od.sd[2]
od.agg <- data.frame(od.mean, od.sd$od.sd)
```

**Figure 2. Mean and SD daily Ontario energy demand (kWh)**
```{r, echo = FALSE, message=FALSE}
od.ts <- ts(od.agg, freq=365, start=c(2002, 5))
plot.ts(od.ts[, c(2:3)], plot.type="single", ylab=NULL, xlab="Date")
```

**Figure 2. Distribtion of Ontario energy demand per hour, 5/2002-3/2004**
```{r, echo = FALSE, message=FALSE}
h <- hist(od$Ontario.Demand, breaks=50, col="gray", main=NULL, xlab="Ontario energy demand (kWh)")
xfit <- seq(min(od$Ontario.Demand), max(od$Ontario.Demand), length=50)
yfit <- dnorm(xfit, mean=mean(od$Ontario.Demand), sd=sd(od$Ontario.Demand))
yfit <- yfit*diff(h$mids[1:2])*length(od$Ontario.Demand) 
lines(xfit, yfit, lwd=2)
```



*2. Calculate the average of each day and plot the standard deviation.*

*3. Use the dataset to forecast (using time series) the daily demand pattern following the steps below:*
*a. Select your hold out sample as the last month of the observations*
*b. Estimate your ARIMA or ARMA model parameters on the rest of your data set*

4. Is it an ARIMA or an ARMA model more suitable? Please justify your choice. What are the
best parameters for the time series model? (10 points)
5. Calculate the predictive accuracy for every day on the test set (in terms of RMSE and
MAPE). (20 points)
6. Can you suggest an alternative method that will be more accurate and flexible to any
potential exogenous shocks in the market? Justify your choice. (10 points)

TEST FOR
TRANSFORMATIONS UNNECESSARY local random variations are consistent over time and generally symmetric in appearance, NO NEED TO STATIONARIZE
