####### Presplit iteration

```{r}
#### 1. Load data 

##### Load base training/test dataset and ensure proper class assignments
```{r}
#train
base.ps = read.csv("C:/Users/Joshua/Google Drive/MSBA/Capstone/Morton/Modeling/Presplit/presplit train.csv")

base.ps$Memid = as.factor(base.ps$Memid)
base.ps$Visits = as.numeric(base.ps$Visits)
base.ps$Rental = as.numeric(base.ps$Rental)
base.ps$Bday_party = as.numeric(base.ps$Bday_party)
base.ps$CnBio.Age = as.numeric(base.ps$CnBio.Age)
base.ps$CnMem.1.01.Consecutive.Years = as.numeric(base.ps$CnMem.1.01.Consecutive.Years)
base.ps = subset(base.ps, select=-c(Expiration.Date))

base.ps$CnMem.1.01.Standing = as.factor(base.ps$CnMem.1.01.Standing)
class(base.ps$CnMem.1.01.Standing)

#imp
imp.ps = read.csv("C:/Users/Joshua/Google Drive/MSBA/Capstone/Morton/Modeling/Presplit/presplit test.csv")

imp.ps$Memid = as.factor(imp.ps$Memid)
imp.ps$Visits = as.numeric(imp.ps$Visits)
imp.ps$Rental = as.numeric(imp.ps$Rental)
imp.ps$Bday_party = as.numeric(imp.ps$Bday_party)
imp.ps$CnBio.Age = as.numeric(imp.ps$CnBio.Age)
imp.ps$CnMem.1.01.Consecutive.Years = as.numeric(imp.ps$CnMem.1.01.Consecutive.Years)
imp.ps = subset(imp.ps, select=-c(Expiration.Date))

imp.ps$CnMem.1.01.Standing = as.factor(imp.ps$CnMem.1.01.Standing)
class(imp.ps$CnMem.1.01.Standing)
```

#### 2. Dealing with missing values

##### Option 1: remove instances with missing values
```{r}
# train
base.ps.sansage = subset(base.ps, select = -c(Age.Group, CnBio.Age))
# Version 1 - excluding age variables, remove the few na instnaces in the Distance variable
base.ps.excludeAge = base.ps.sansage[complete.cases(base.ps.sansage),]
# version 2 - including age variables, remove all na instnaces (approx 50%)
base.ps.includeAge = base.ps[complete.cases(base.ps),]
# remove age and age group
imp.ps.sansage = subset(imp.ps, select = -c(Age.Group, CnBio.Age))

# imp
# Version 1 - excluding age variables, remove the few na instnaces in the Distance variable
imp.ps.excludeAge = imp.ps.sansage[complete.cases(imp.ps.sansage),]
# version 2 - including age variables, remove all na instnaces (approx 50%)
imp.ps.includeAge = imp.ps[complete.cases(imp.ps),]
```

##### remove Memid 
```{r}
# train
base.ps.excludeAge = subset(base.ps.excludeAge, select=-c(Memid))
base.ps.includeAge = subset(base.ps.includeAge, select=-c(Memid))

# imp
imp.ps.excludeAge = subset(imp.ps.excludeAge, select=-c(Memid))
imp.ps.includeAge = subset(imp.ps.includeAge, select=-c(Memid))

```

> Accounting for missing values has resulted in the following data sets:

 - Remove missing values from training set
   - base.ps.excludeAge: excluding age variables and removing all na instances
   - base.ps.includeAge: including age variables and removing all na instances

 - Remove missing values from implementation set
   - imp.ps.excludeAge: excluding age variables and removing all na instances
   - imp.ps.includeAge: including age variables and removing all na instances

#### 4. Apply transformations (boxcox, center, scale, pca)
```{r}
# train
# first apply log transformations to numeric variables - only applied to distance as all other numeric variables contain a significant amount of zeros

base.ps.excludeAge$Distance.log = log(base.ps.excludeAge$Distance)
base.ps.includeAge$Distance.log = log(base.ps.includeAge$Distance)

numerics.age2 = c("Visits", "NetAmt", "Rental", "Bday_party", "Events", "CnBio.Age", "CnMem.1.01.Consecutive.Years", "Distance", "Distance.log")
numerics.noage2 = c("Visits", "NetAmt", "Rental", "Bday_party", "Events", "CnMem.1.01.Consecutive.Years", "Distance", "Distance.log")

# reassess for skewnesss
base.ps.excludeAge.N2 = base.ps.excludeAge[,numerics.noage2]
base.ps.includeAge.N2 = base.ps.includeAge[,numerics.age2]

# ignores categorical variables and only applies the preProcess function to numeric varaibles
# essentially replace preProcess with ppWrapper
# saves results as dataframe
ppWrapper = function( x, excludeClasses=c("factor"), ... ) {
    whichToExclude <- sapply( x, function(y) any(sapply(excludeClasses, function(excludeClass) is(y,excludeClass) )) )
    processedMat <- predict( preProcess( x[!whichToExclude], ...), newdata=x[!whichToExclude] )
    x[!whichToExclude] <- processedMat
    x
}

library(caret)
base.ps.excludeAge.Proc = ppWrapper(base.ps.excludeAge, method = c("BoxCox", "center", "scale","pca"))                                                             
base.ps.includeAge.Proc = ppWrapper(base.ps.includeAge, method = c("BoxCox", "center", "scale","pca"))

# remove distance since distance.log created
base.ps.excludeAge.Proc2 = subset(base.ps.excludeAge.Proc, select=-c(Distance))                                                           
base.ps.includeAge.Proc2 = subset(base.ps.includeAge.Proc, select=-c(Distance))  

#imp

# first apply log transformations to numeric variables - only applied to distance as all other numeric variables contain a significant amount of zeros

imp.ps.excludeAge$Distance.log = log(imp.ps.excludeAge$Distance)
imp.ps.includeAge$Distance.log = log(imp.ps.includeAge$Distance)

# reassess for skewnesss
imp.ps.excludeAge.N2 = imp.ps.excludeAge[,numerics.noage2]
imp.ps.includeAge.N2 = imp.ps.includeAge[,numerics.age2]

imp.ps.excludeAge.Proc = ppWrapper(imp.ps.excludeAge, method = c("BoxCox", "center", "scale","pca"))                                                             
imp.ps.includeAge.Proc = ppWrapper(imp.ps.includeAge, method = c("BoxCox", "center", "scale","pca"))

# remove distance since distance.log created
imp.ps.excludeAge.Proc2 = subset(imp.ps.excludeAge.Proc, select=-c(Distance))                                                           
imp.ps.includeAge.Proc2 = subset(imp.ps.includeAge.Proc, select=-c(Distance))  

```

> Transformations lowered skewness values. Created the following files with transformed numeric variables and original nominal variables:
   - base.ps.excludeAge.Proc2                                                             
   - base.ps.includeAge.Proc2

> Transformations lowered skewness values. Created the following files with transformed numeric variables and original nominal variables:
   - imp.ps.excludeAge.Proc2                                                             
   - imp.ps.includeAge.Proc2
   
#### 3. Convert all categorical/nominal variables to numeric Dummy variables
```{r}

# train

base.ps.excludeAge.Proc2$CnMem.1.01.Standing = as.numeric(base.ps.excludeAge.Proc2$CnMem.1.01.Standing) #convert target to numeric to prevent dummification
a.ps = dummyVars("~ .", data=base.ps.excludeAge.Proc2, levelsOny=TRUE)
base.ps.excludeAge.Proc.dummy = data.frame(predict(a.ps, newdata = base.ps.excludeAge.Proc2))
base.ps.excludeAge.Proc.dummy$CnMem.1.01.Standing = as.factor(base.ps.excludeAge.Proc.dummy$CnMem.1.01.Standing)# convert back to factor
base.ps.excludeAge.Proc.dummy$CnMem.1.01.Standing = as.factor(ifelse(base.ps.excludeAge.Proc.dummy$CnMem.1.01.Standing == "1", "0", "1"))

base.ps.includeAge.Proc2$CnMem.1.01.Standing = as.numeric(base.ps.includeAge.Proc2$CnMem.1.01.Standing)
b.ps = dummyVars("~ .", data=base.ps.includeAge.Proc2, levelsOny=TRUE)
base.ps.includeAge.Proc.dummy = data.frame(predict(b.ps, newdata = base.ps.includeAge.Proc2))
base.ps.includeAge.Proc.dummy$CnMem.1.01.Standing = as.factor(base.ps.includeAge.Proc.dummy$CnMem.1.01.Standing)
base.ps.includeAge.Proc.dummy$CnMem.1.01.Standing = as.factor(ifelse(base.ps.includeAge.Proc.dummy$CnMem.1.01.Standing == "1", "0", "1"))
#imp

imp.ps.excludeAge.Proc2$CnMem.1.01.Standing = as.numeric(imp.ps.excludeAge.Proc2$CnMem.1.01.Standing) #convert target to numeric to prevent dummification

a.ps.i = dummyVars("~ .", data=imp.ps.excludeAge.Proc2, levelsOny=TRUE)
imp.ps.excludeAge.Proc.dummy = data.frame(predict(a.ps.i, newdata = imp.ps.excludeAge.Proc2))
imp.ps.excludeAge.Proc.dummy$CnMem.1.01.Standing = as.factor(imp.ps.excludeAge.Proc.dummy$CnMem.1.01.Standing)# convert back to factor
imp.ps.excludeAge.Proc.dummy$CnMem.1.01.Standing = as.factor(ifelse(imp.ps.excludeAge.Proc.dummy$CnMem.1.01.Standing == "1", "0", "1"))

imp.ps.includeAge.Proc2$CnMem.1.01.Standing = as.numeric(imp.ps.includeAge.Proc2$CnMem.1.01.Standing)

b.ps.i = dummyVars("~ .", data=imp.ps.includeAge.Proc2, levelsOny=TRUE)
imp.ps.includeAge.Proc.dummy = data.frame(predict(b.ps.i, newdata = imp.ps.includeAge.Proc2))
imp.ps.includeAge.Proc.dummy$CnMem.1.01.Standing = as.factor(imp.ps.includeAge.Proc.dummy$CnMem.1.01.Standing)
imp.ps.includeAge.Proc.dummy$CnMem.1.01.Standing = as.factor(ifelse(imp.ps.includeAge.Proc.dummy$CnMem.1.01.Standing == "1", "0", "1"))

# rerun to restore pre-dummy sets
base.ps.excludeAge.Proc2 = subset(base.ps.excludeAge.Proc, select=-c(Distance))                                                           
base.ps.includeAge.Proc2 = subset(base.ps.includeAge.Proc, select=-c(Distance))  
imp.ps.excludeAge.Proc2 = subset(imp.ps.excludeAge.Proc, select=-c(Distance))                                                           
imp.ps.includeAge.Proc2 = subset(imp.ps.includeAge.Proc, select=-c(Distance))  

```

> Created the following files with transformed numeric variables and dummified nominal variables:
   - base.ps.excludeAge.Proc.dummy                                                             
   - base.ps.includeAge.Proc.dummy

> Created the following files with transformed numeric variables and dummified nominal variables:
   - imp.ps.excludeAge.Proc.dummy                                                             
   - imp.ps.includeAge.Proc.dummy
   
   
#### 4. Train and eval model on preprocessed data sets
```{r}
library(pROC)
library(RWeka)
library(randomForest)
library(C50)


# remove membership types from training set that are not in imp data set
str(base.ps.excludeAge.Proc.dummy)
str(imp.ps.excludeAge.Proc.dummy)
base.ps.excludeAge.Proc.dummy.rem = subset(base.ps.excludeAge.Proc.dummy, select = -c(CnMem.1.01.Cur.Category.description.2.YR.Family.Plus, CnMem.1.01.Cur.Category.description.Ginkgo.Circle, CnMem.1.01.Cur.Category.description.Retiree, CnMem.1.01.Cur.Category.description.Trustee,  CnBio.Constit.Code.Trustee ))

#run j48
J48.TRN.BpEPD = J48(CnMem.1.01.Standing ~ ., base.ps.excludeAge.Proc.dummy.rem)
J48.TST.BpEPD = predict(J48.TRN.BpEPD, imp.ps.excludeAge.Proc.dummy, type="prob")

imp.ps.excludeAge.Proc.dummy$J48prob = J48.TST.BpEPD[,"1"]
J48.BpEPD.ROC = roc(response = imp.ps.excludeAge.Proc.dummy$CnMem.1.01.Standing, predictor = imp.ps.excludeAge.Proc.dummy$J48prob, levels = rev(levels(imp.ps.excludeAge.Proc.dummy$CnMem.1.01.Standing)))

auc(J48.BpEPD.ROC)
plot(J48.BpEPD.ROC, legacy.axes = TRUE)

#run LMT
LMT.TRN.BpEPD = LMT(CnMem.1.01.Standing ~ ., base.ps.excludeAge.Proc.dummy.rem)
LMT.TST.BpEPD = predict(LMT.TRN.BpEPD, imp.ps.excludeAge.Proc.dummy, type="prob")

imp.ps.excludeAge.Proc.dummy$LMTprob = LMT.TST.BpEPD[,"1"]
LMT.BpEPD.ROC = roc(response = imp.ps.excludeAge.Proc.dummy$CnMem.1.01.Standing, predictor = imp.ps.excludeAge.Proc.dummy$LMTprob, levels = rev(levels(imp.ps.excludeAge.Proc.dummy$CnMem.1.01.Standing)))

auc(LMT.BpEPD.ROC)
plot(LMT.BpEPD.ROC, legacy.axes = TRUE)

#run randomForest
RF.TRN.BpEPD = randomForest(CnMem.1.01.Standing ~ ., base.ps.excludeAge.Proc.dummy.rem)
RF.TST.BpEPD = predict(RF.TRN.BpEPD, imp.ps.excludeAge.Proc.dummy, type="prob")

imp.ps.excludeAge.Proc.dummy$RFprob = RF.TST.BpEPD[,"1"]
RF.BpEPD.ROC = roc(response = imp.ps.excludeAge.Proc.dummy$CnMem.1.01.Standing, predictor = imp.ps.excludeAge.Proc.dummy$RFprob, levels = rev(levels(imp.ps.excludeAge.Proc.dummy$CnMem.1.01.Standing)))

auc(RF.BpEPD.ROC)
plot(RF.BpEPD.ROC, legacy.axes = TRUE)

#run C5.0
C5.0.TRN.BpEPD = C5.0(CnMem.1.01.Standing ~ ., base.ps.excludeAge.Proc.dummy.rem)
C5.0.TST.BpEPD = predict(C5.0.TRN.BpEPD, imp.ps.excludeAge.Proc.dummy, type="prob")

imp.ps.excludeAge.Proc.dummy$C5.0prob = C5.0.TST.BpEPD[,"1"]
C5.0.BpEPD.ROC = roc(response = imp.ps.excludeAge.Proc.dummy$CnMem.1.01.Standing, predictor = imp.ps.excludeAge.Proc.dummy$C5.0prob, levels = rev(levels(imp.ps.excludeAge.Proc.dummy$CnMem.1.01.Standing)))

auc(C5.0.BpEPD.ROC)
plot(C5.0.BpEPD.ROC, legacy.axes = TRUE)


```

#### 5. Train and eval model on un-preprocessed (except missing values) data sets
 
```{r}

# align nominal predictor levels (i.e. remove levels/instances from training set not in test set)

str(base.ps.excludeAge)
str(imp.ps.excludeAge)
# CnBio.Constit.Code
levels(imp.ps.excludeAge$CnBio.Constit.Code)
levels(base.ps.excludeAge$CnBio.Constit.Code) # remove Trustee
levels(base.ps.excludeAge$CnMem.1.01.Cur.Category.description)
levels(imp.ps.excludeAge$CnMem.1.01.Cur.Category.description) # remove 2 YR Family Plus, Ginkgo Circle, Retiree, Trustee

base.ps.excludeAge.rem = base.ps.excludeAge[!(base.ps.excludeAge$CnBio.Constit.Code %in% c("Trustee") | base.ps.excludeAge$CnMem.1.01.Cur.Category.description %in% c("2 YR Family Plus", "Ginkgo Circle", "Retiree", "Trustee")),]

# excludeage dataset
# j48
J48.TRN.BpE = J48(CnMem.1.01.Standing ~ ., base.ps.excludeAge.rem) # train on test set
J48.TST.BpE = predict(J48.TRN.BpE, imp.ps.excludeAge, type="prob") # apply model on holdout set to get pred class and class probs

J48prob = J48.TST.BpE[,"1"] #grab the positive class, i.e. 1/churn
J48.BpE.ROC = roc(response = imp.ps.excludeAge$CnMem.1.01.Standing, predictor = J48prob)
confusionMatrix(data = J48prob, reference = imp.ps.excludeAge$CnMem.1.01.Standing, positive="1")
auc(J48.BpE.ROC)
plot(J48.BpE.ROC, legacy.axes = TRUE)

#j48 - weka verification
BpE.eval_J48 = evaluate_Weka_classifier(J48.TRN.BpE, imp.ps.excludeAge, class=TRUE)
BpE.eval_J48



# includeage dataset
# j48
J48.TRN.BpI = J48(CnMem.1.01.Standing ~ ., base.ps.includeAge) # train on test set
J48.TST.BpI = predict(J48.TRN.BpI, imp.ps.includeAge, type="prob") # apply model on holdout set

J48prob.I = J48.TST.BpI[,"1"]
J48.BpI.ROC = roc(response = imp.ps.includeAge$CnMem.1.01.Standing, predictor = J48prob.I)
confusionMatrix(data = J48prob.I, reference = imp.ps.includeAge$CnMem.1.01.Standing, positive="1")
auc(J48.BpI.ROC)
plot(J48.BpI.ROC, legacy.axes = TRUE)

#j48 - weka verification
BpI.eval_J48 = evaluate_Weka_classifier(J48.TRN.BpI, imp.ps.includeAge, class=TRUE)
BpI.eval_J48





